HFI in RISC-V

### 2025-05-01 paper.tex
h-prefix
Cleanup paragraph formatting

\section{Implementation \& Evaluation}
\subsection{Implementation - Intial Setup}
Since we were implenting our RISC-V HFI extension in QEMU, we cloned the current open-source QEMU repository and began work there. We began by taking time to understand how QEMU works, the purpose of the various repositories, and gaining a better understanding of RISC-V. We worked with the assumption that the system being run on always supports HFI but it is an easy addition to add any necessary hardware checks to validate that assumption.\\
Our work then began by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions. Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
In order to implement our instruction we have opted to use the free opcode $0001011$.
We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$. Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.\\

\subsection{Implementation - Implicit Regions}
After testing whether we can toggle the \textit{hfi\_status} register properly, we moved on to implement implicit regions. As aforementioned, implicit regions are used for native, unmodified binaries and an important feature of HFI. The original HFI paper had four implicit data regions and two implicit code regions. In our implementation, we supported an arbitrary number of data and code regions which were defined by a global variable that could be adjusted as needed. We also tested using varying numbers of each region and ensured the HFI logic worked seamlessly with the defined number of regions. Implicit regions are defined by a base and a mask.\\
To store metadata about these regions, we created new structs \textit{HFIImplicitDataRegion} and \textit{HFIImplicitCodeRegion} which we stored in the CPU architecture state in arrays of length based on the number of regions we were supporting. To set these regions, we implemented the instruction \textit{hfi\_set\_region\_size} which took in a region number, base address, and mask and configured a new data or code region accordingly. Then, to set permissions for this region, we added a new instruction \textit{hfi\_set\_region\_permissions} which took a region number and a 64 bit integer encoding the permissions for that region. Since a single call to \textit{hfi\_set\_region\_permissions} will be used for a singular region (differentiated using the code region number), the permission encoding for each of the 3 types of region can also differ. Implicit data regions store the enabled bit at bit 7, read bit at bit 6, and write bit at bit 5. Implicit code regions store the enabled bit at bit 7 and executable bit at bit 6.\\
Although implicit regions are used for unmodified native binaries, we are still working with the assumption that the software setting up HFI for the binary is a trustable user who will setup the regions properly. However, one check we do insert is for uninitialized regions. Originally, if a region wasn't setup properly where \textit{hfi\_set\_region\_size} wasn't called but \textit{hfi\_set\_region\_permissions} was called, a region would still be configured with $0x0$ for the base and mask. This was an issue as such a region allowed any code/data addresses to pass through. Hence, to prevent this, we set the mask on implicit regions to $0x11111111$ initially which would prevent regions not set up properly to fail HFI checks.\\
To implement our instruction we have opted to use the free opcode $0001011$.
We are then able to differentiate between these functions and the two aforementioned functions (\textit{hfi\_enter} and \textit{hfi\_exit}) by modifying the funct3 value of the instructions such that \textit{hfi\_set\_region\_size} has funct3 $= 010$ and \textit{hfi\_set\_region\_permissions} has funct3 $= 011$.

\subsection{Implementation - Explicit Regions}
Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExpllicitDataRegion} to store metadata about explicit regions. These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support. While the original HFI paper had the h-prefix \textit{mov} instruct (\textit{hmov}), we had h-prefix load and store instruction. This difference is due to the ISA differences between x86 and RISC-V. Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.\\
For each of the load and store instructions, we would check the whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions). If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction. If the check fails, HFI would trap and prevent access to the address.\\
Explicit data regions are configured using a base and a bound. To set these regions, we use the same \textit{hfi\_set\_region\_base} and \textit{hfi\_set\_region\_permissions} that we used for implicit regions. The value passed for a mask in implicit regions was instead used as the bound value when an explicit region was configured. For the permissions, explicit data regions store the enabled bit at bit 7, read bit at bit 6, write bit at bit 5, and large region bit at bit 4.\\
We tested the implementation using this single explicit data region. One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in section 4. However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in section 6 can allow easy extension to support more explicit data regions.\\
For these new load instructions, we used the free opcode $0001011$. For the new store instructions, we used the free opcode $1011011$. You can find the funct3 values we used to differentiate between them in Table~\ref{tab:hfi_instructions}.

\begin{table}[]
\caption{Encodings of HFI instruction implemented}
\label{tab:hfi_instructions}
\begin{tabular}{|l|l|l|}
\hline
                                         &  opcode   & funct3   \\ \hline
\textit{hfi\_enter}                      &  0001011  &  000     \\ \hline
\textit{hfi\_exit}                       &  0001011  &  001     \\ \hline
\textit{hfi\_set\_region\_size}          &  0001011  &  010     \\ \hline
\textit{hfi\_set\_region\_permissions}   &  0001011  &  011     \\ \hline
\textit{hfi\_hlb0}                       &  0101011  &  000     \\ \hline
\textit{hfi\_hlh0}                       &  0101011  &  001     \\ \hline
\textit{hfi\_hlw0}                       &  0101011  &  010     \\ \hline
\textit{hfi\_hlbu0}                      &  0101011  &  100     \\ \hline
\textit{hfi\_hlhu0}                      &  0101011  &  101     \\ \hline
\textit{hfi\_hlwu0}                      &  0101011  &  110     \\ \hline
\textit{hfi\_hld0}                       &  0101011  &  011     \\ \hline
\textit{hfi\_hsb0}                       &  1011011  &  010     \\ \hline
\textit{hfi\_hsh0}                       &  1011011  &  011     \\ \hline
\textit{hfi\_hsw0}                       &  1011011  &  100     \\ \hline
\textit{hfi\_hsd0}                       &  1011011  &  000     \\ \hline
\end{tabular}
\end{table}

\subsection{Evaluation - Testing}
To validate our HFI implementation and ensure correct functionality across all features, we developed a comprehensive testing infrastructure. Our testing framework consists of a suite of assembly programs written specifically to test different aspects of HFI functionality on RISC-V. These tests are compiled using the RISC-V GCC toolchain and executed on our modified QEMU binary.\\
We use a modular testing approach where each assembly program focuses on testing a specific component or feature of HFI. Our test suite covers several key areas of functionality: basic HFI operations that verify the core instructions function correctly and properly toggle the HFI status; region configuration tests that ensure memory regions can be properly defined with appropriate size and permissions; memory access tests that verify operations within configured regions succeed while operations outside them fail appropriately; and explicit region access tests that exercise the h-prefix load/store instructions to validate boundary checking and permission enforcement. We also implemented control tests to establish baseline functionality in non-HFI contexts.\\
These test programs are written in RISC-V assembly and leverage a custom set of macros defined in our includes directory that encode our HFI instructions using the appropriate opcodes and operands. For example, our HFI enter macro encodes the instruction with opcode 0x0B and function code 0 along with the specified register operands. This approach allows us to write readable assembly that directly tests our custom instructions.\\
We developed a Makefile-based build system that compiles these assembly programs using the RISC-V toolchain (riscv64-unknown-elf-as and riscv64-unknown-elf-ld). The resulting ELF binaries are then executed on our custom-built QEMU with additional debug flags to capture instruction execution and CPU state information.\\
The test execution outputs are captured in log files, allowing us to analyze instruction execution, register state changes, and any traps or exceptions that occur during testing. This approach enables us to verify several critical aspects of the implementation: the HFI instructions execute with the expected opcodes, region configurations are properly stored in CPU state, memory accesses are correctly validated against region bounds, and violations trigger appropriate traps when they should occur.\\
Through this testing infrastructure, we confirmed the correctness of our implementation across various scenarios including successful sandboxing, proper isolation of memory regions, and correct handling of boundary conditions. The modular nature of our test suite allows us to incrementally validate features as they are implemented and regression-test them as the implementation evolves.

@@

\section{Related Work}
Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}. HFI provides a simple ISA extension to support secure, flexible, and efficient in-process isolation, addressing limitations of existing software-based techniques, including runtime overheads, scalability limits, Spectre vulnerability, and compatibility issues.\\
There are also several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques. Some systems have proposed extensions to the page table metadata to store a per-sandbox ID checked by hardware, such as Donky \cite{Donky} or IMIX \cite{IMIX}. Donky is a hardware-software co-design providing strong in-process isolation based on dynamic memory protection domains using hardware-backed protection keys. Its implementation on RISC-V uses the reserved top 10 bits of page-table entries for protection keys, supporting 1024 keys per process. Donky distinguishes itself by managing protection key policies entirely in userspace via a dedicated control and status register (DKRU). Access to the DKRU is protected by a hardware call gate, which leverages the RISC-V N extension. This architecture provides substantially stronger security guarantees than systems like Intel MPK, where the policy register is unprivileged, allowing Donky to shield against arbitrary code execution without needing binary inspection or rewriting. Donky facilitates fast domain switches entirely in userspace, with minimal hardware additions on RISC-V. Donky has been implemented for both x86 (using MPK emulation) and RISC-V architectures, while IMIX has only been implemented on x86.\\
IMIX (In-Process Memory Isolation Extension) introduces a minimal x86 ISA extension for isolating sensitive data within a process \cite{IMIX}. Unlike many isolation mechanisms tied to a specific defense (e.g., MPK for protection keys or CET for shadow stacks), IMIX provides a mitigation-agnostic hardware primitive for protecting security-critical data. IMIX enables memory pages to be marked as isolated using a new permission flag. These pages can only be accessed with a special instruction, smov, introduced by IMIX. This mechanism allows a wide range of memory-corruption defenses (such as CPI or CFI) to protect their internal metadata from arbitrary access by potentially compromised application code. IMIX was prototyped using Intel’s simulation infrastructure, with extensions to LLVM and Linux to support its new instruction semantics. The authors demonstrated that IMIX significantly reduces the attack surface for in-process defenses with low overhead and better flexibility than existing solutions like MPK, which are limited by unprivileged key registers and require complex mitigation workarounds to be secure.\\
In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in CHERI \cite{CHERI}. While initially developed as a hybrid capability model extending the 64-bit MIPS ISA and demonstrated on FPGA, the CHERI approach is described as not being specific to MIPS and is a prominent capability system relevant to discussions of advanced ISAs, including RISC-V (as referenced in related work comparisons). CHERI provides byte-granularity memory protection and enables hardware enforcement of memory safety and fault isolation. It utilizes a capability coprocessor and tagged memory, where capabilities are stored in dedicated registers and memory locations, protected by hardware tags and unforgeable manipulation instructions. Capabilities are unforgeable and define protection domains based on reachability. CHERI aims for a high level of source-code and binary compatibility relative to prior pure capability machines, supporting incremental deployment. However, adapting and recompiling application code is generally required to utilize its full protection features, and achieving comprehensive object-level memory safety involves extensive modifications throughout the software stack (OS, ABI, compiler). CHERI offers strong fine-grained protection within an address space and supports scalable userspace protection domains.\\

>>>>

\section{Implementation \& Evaluation}
\subsection{Implementation - Intial Setup}
Since we were implenting our RISC-V HFI extension in QEMU, we cloned the current open-source QEMU repository and began work there. We began by taking time to understand how QEMU works, the purpose of the various repositories, and gaining a better understanding of RISC-V. We worked with the assumption that the system being run on always supports HFI but it is an easy addition to add any necessary hardware checks to validate that assumption.

Our work then began by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions. Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
In order to implement our instruction we have opted to use the free opcode $0001011$.
We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$. Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.

\subsection{Implementation - Implicit Regions}
After testing whether we can toggle the \textit{hfi\_status} register properly, we moved on to implement implicit regions. As aforementioned, implicit regions are used for native, unmodified binaries and an important feature of HFI. The original HFI paper had four implicit data regions and two implicit code regions. In our implementation, we supported an arbitrary number of data and code regions which were defined by a global variable that could be adjusted as needed. We also tested using varying numbers of each region and ensured the HFI logic worked seamlessly with the defined number of regions. Implicit regions are defined by a base and a mask.

To store metadata about these regions, we created new structs \textit{HFIImplicitDataRegion} and \textit{HFIImplicitCodeRegion} which we stored in the CPU architecture state in arrays of length based on the number of regions we were supporting. To set these regions, we implemented the instruction \textit{hfi\_set\_region\_size} which took in a region number, base address, and mask and configured a new data or code region accordingly. Then, to set permissions for this region, we added a new instruction \textit{hfi\_set\_region\_permissions} which took a region number and a 64 bit integer encoding the permissions for that region. Since a single call to \textit{hfi\_set\_region\_permissions} will be used for a singular region (differentiated using the code region number), the permission encoding for each of the 3 types of region can also differ. Implicit data regions store the enabled bit at bit 7, read bit at bit 6, and write bit at bit 5. Implicit code regions store the enabled bit at bit 7 and executable bit at bit 6.

Although implicit regions are used for unmodified native binaries, we are still working with the assumption that the software setting up HFI for the binary is a trustable user who will setup the regions properly. However, one check we do insert is for uninitialized regions. Originally, if a region wasn't setup properly where \textit{hfi\_set\_region\_size} wasn't called but \textit{hfi\_set\_region\_permissions} was called, a region would still be configured with $0x0$ for the base and mask. This was an issue as such a region allowed any code/data addresses to pass through. Hence, to prevent this, we set the mask on implicit regions to $0x11111111$ initially which would prevent regions not set up properly to fail HFI checks.

To implement our instruction we have opted to use the free opcode $0001011$.
We are then able to differentiate between these functions and the two aforementioned functions (\textit{hfi\_enter} and \textit{hfi\_exit}) by modifying the funct3 value of the instructions such that \textit{hfi\_set\_region\_size} has funct3 $= 010$ and \textit{hfi\_set\_region\_permissions} has funct3 $= 011$.

\subsection{Implementation - Explicit Regions}
Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExpllicitDataRegion} to store metadata about explicit regions. These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support. While the original HFI paper had the h-prefix \textit{mov} instruct (\textit{hmov}), we had h-prefix load and store instruction. This difference is due to the ISA differences between x86 and RISC-V. Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.

For each of the load and store instructions, we would check the whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions). If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction. If the check fails, HFI would trap and prevent access to the address.

Explicit data regions are configured using a base and a bound. To set these regions, we use the same \textit{hfi\_set\_region\_base} and \textit{hfi\_set\_region\_permissions} that we used for implicit regions. The value passed for a mask in implicit regions was instead used as the bound value when an explicit region was configured. For the permissions, explicit data regions store the enabled bit at bit 7, read bit at bit 6, write bit at bit 5, and large region bit at bit 4.

We tested the implementation using this single explicit data region. One obvious limitation is how we can scale this to work for multiple data regions, which can be a challenge due to the opcode limitation discussed in section 4. However, based on the feasibility of our implementation for the one data region and overall structure of the implementation, we believe proposed ideas in section 6 can allow easy extension to support more explicit data regions.

For these new load instructions, we used the free opcode $0001011$. For the new store instructions, we used the free opcode $1011011$. You can find the funct3 values we used to differentiate between them in Table~\ref{tab:hfi_instructions}.

\begin{table}[]
\caption{Encodings of HFI instruction implemented}
\label{tab:hfi_instructions}
\begin{tabular}{|l|l|l|}
\hline
                                         &  opcode   & funct3   \\ \hline
\textit{hfi\_enter}                      &  0001011  &  000     \\ \hline
\textit{hfi\_exit}                       &  0001011  &  001     \\ \hline
\textit{hfi\_set\_region\_size}          &  0001011  &  010     \\ \hline
\textit{hfi\_set\_region\_permissions}   &  0001011  &  011     \\ \hline
\textit{hlb0}                       &  0101011  &  000     \\ \hline
\textit{hlh0}                       &  0101011  &  001     \\ \hline
\textit{hlw0}                       &  0101011  &  010     \\ \hline
\textit{hlbu0}                      &  0101011  &  100     \\ \hline
\textit{hlhu0}                      &  0101011  &  101     \\ \hline
\textit{hlwu0}                      &  0101011  &  110     \\ \hline
\textit{hld0}                       &  0101011  &  011     \\ \hline
\textit{hsb0}                       &  1011011  &  010     \\ \hline
\textit{hsh0}                       &  1011011  &  011     \\ \hline
\textit{hsw0}                       &  1011011  &  100     \\ \hline
\textit{hsd0}                       &  1011011  &  000     \\ \hline
\end{tabular}
\end{table}

\subsection{Evaluation - Testing}
To validate our HFI implementation and ensure correct functionality across all features, we developed a comprehensive testing infrastructure. Our testing framework consists of a suite of assembly programs written specifically to test different aspects of HFI functionality on RISC-V. These tests are compiled using the RISC-V GCC toolchain and executed on our modified QEMU binary.

We use a modular testing approach where each assembly program focuses on testing a specific component or feature of HFI. Our test suite covers several key areas of functionality: basic HFI operations that verify the core instructions function correctly and properly toggle the HFI status; region configuration tests that ensure memory regions can be properly defined with appropriate size and permissions; memory access tests that verify operations within configured regions succeed while operations outside them fail appropriately; and explicit region access tests that exercise the h-prefix load/store instructions to validate boundary checking and permission enforcement. We also implemented control tests to establish baseline functionality in non-HFI contexts.

These test programs are written in RISC-V assembly and leverage a custom set of macros defined in our includes directory that encode our HFI instructions using the appropriate opcodes and operands. For example, our HFI enter macro encodes the instruction with opcode 0x0B and function code 0 along with the specified register operands. This approach allows us to write readable assembly that directly tests our custom instructions.

We developed a Makefile-based build system that compiles these assembly programs using the RISC-V toolchain (riscv64-unknown-elf-as and riscv64-unknown-elf-ld). The resulting ELF binaries are then executed on our custom-built QEMU with additional debug flags to capture instruction execution and CPU state information.

The test execution outputs are captured in log files, allowing us to analyze instruction execution, register state changes, and any traps or exceptions that occur during testing. This approach enables us to verify several critical aspects of the implementation: the HFI instructions execute with the expected opcodes, region configurations are properly stored in CPU state, memory accesses are correctly validated against region bounds, and violations trigger appropriate traps when they should occur.

Through this testing infrastructure, we confirmed the correctness of our implementation across various scenarios including successful sandboxing, proper isolation of memory regions, and correct handling of boundary conditions. The modular nature of our test suite allows us to incrementally validate features as they are implemented and regression-test them as the implementation evolves.

@@

\section{Related Work}
Our work on HFI for RSIC-V builds on prior work that established HFI for x86-64 architectures \cite{HFI}. HFI provides a simple ISA extension to support secure, flexible, and efficient in-process isolation, addressing limitations of existing software-based techniques, including runtime overheads, scalability limits, Spectre vulnerability, and compatibility issues.

There are also several alternate approaches have been developed over the years for fine-grained isolation involving hardware-assisted techniques. Some systems have proposed extensions to the page table metadata to store a per-sandbox ID checked by hardware, such as Donky \cite{Donky} or IMIX \cite{IMIX}. Donky is a hardware-software co-design providing strong in-process isolation based on dynamic memory protection domains using hardware-backed protection keys. Its implementation on RISC-V uses the reserved top 10 bits of page-table entries for protection keys, supporting 1024 keys per process. Donky distinguishes itself by managing protection key policies entirely in userspace via a dedicated control and status register (DKRU). Access to the DKRU is protected by a hardware call gate, which leverages the RISC-V N extension. This architecture provides substantially stronger security guarantees than systems like Intel MPK, where the policy register is unprivileged, allowing Donky to shield against arbitrary code execution without needing binary inspection or rewriting. Donky facilitates fast domain switches entirely in userspace, with minimal hardware additions on RISC-V. Donky has been implemented for both x86 (using MPK emulation) and RISC-V architectures, while IMIX has only been implemented on x86.

IMIX (In-Process Memory Isolation Extension) introduces a minimal x86 ISA extension for isolating sensitive data within a process \cite{IMIX}. Unlike many isolation mechanisms tied to a specific defense (e.g., MPK for protection keys or CET for shadow stacks), IMIX provides a mitigation-agnostic hardware primitive for protecting security-critical data. IMIX enables memory pages to be marked as isolated using a new permission flag. These pages can only be accessed with a special instruction, smov, introduced by IMIX. This mechanism allows a wide range of memory-corruption defenses (such as CPI or CFI) to protect their internal metadata from arbitrary access by potentially compromised application code. IMIX was prototyped using Intel’s simulation infrastructure, with extensions to LLVM and Linux to support its new instruction semantics. The authors demonstrated that IMIX significantly reduces the attack surface for in-process defenses with low overhead and better flexibility than existing solutions like MPK, which are limited by unprivileged key registers and require complex mitigation workarounds to be secure.

In addition to this, there are also methods that make use of capability-based addressing, a scheme for controlling access to memory, which is seen most prominently in CHERI \cite{CHERI}. While initially developed as a hybrid capability model extending the 64-bit MIPS ISA and demonstrated on FPGA, the CHERI approach is described as not being specific to MIPS and is a prominent capability system relevant to discussions of advanced ISAs, including RISC-V (as referenced in related work comparisons). CHERI provides byte-granularity memory protection and enables hardware enforcement of memory safety and fault isolation. It utilizes a capability coprocessor and tagged memory, where capabilities are stored in dedicated registers and memory locations, protected by hardware tags and unforgeable manipulation instructions. Capabilities are unforgeable and define protection domains based on reachability. CHERI aims for a high level of source-code and binary compatibility relative to prior pure capability machines, supporting incremental deployment. However, adapting and recompiling application code is generally required to utilize its full protection features, and achieving comprehensive object-level memory safety involves extensive modifications throughout the software stack (OS, ABI, compiler). CHERI offers strong fine-grained protection within an address space and supports scalable userspace protection domains.
